Notes of Basic Topolgy

### 2016-12-27 Basic Topology Notes/sections/ch5-fundamental-groups.tex
[major] new note for topology: simplicial approximation theorem
[major] Talks about stellar subdivision and how homology group are a homotopic invariant.

\paragraph{The $n$-sphere $S^n$} is on page 99. To prove it, we
require a theorem:
\begin{thm}
    Let $X$ be a space which can be written as the union of two simply
    connected open sets $U,V$ in such a way that $U\cap V$ is
    path-connected. Then $X$ is simply connected.
\end{thm}
\begin{proof}
    Proof is on page 99 of \cite{book}. It requires the Lebesgue lemma
    \ref{thm:lebesgue-lemma}. A picture illustrating the proof is
    provided here:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]
        {pics/n-sphere-is-simply-con.PNG}
    \end{figure}
\end{proof}
Then, the two cover of $S^n$ are simply $U=S^n-\{x\}$ and
$V=S^n-\{y\}$, where $x,y$ are two distinct points on $S^n$. Since $U$
and $V$ are homeomorphic to $\E^n$, $S^n$ is simply-connected.

@@

We have an application,
\begin{thm}[Brouwer Fixed Point theorem]
    A contiuous map $f:B^2\to B^2$ ($B^2:2D$-closed dicks) must have a
fixed point. That is, $\exists x\in B^2$ such that $f(x)=x$.
\end{thm}
\begin{proof}
    Assuming that this theorem is false, that is $\forall x\in B^2$,
    $x\neq f(x)$, then we have a straight path from $f(x)$ to $x$. We can
    extends this path to cuts the boundary of $B^2$ at $h(x)$. This is
    for all $x\in B^2$, hence we have a map $h: B^2 \to S^1$. Also,
    $h|_{S^1}$ is obvious an identity map. But $S^1$ can be included
    inside $B^2$, so we have:
    \begin{equation}
        S^1 \to B^2 \to S^1
    \end{equation}
    Hence we have a series of homomorphism of fundamental groups:
    \begin{equation}
        \pi_1 (S^1) \to \pi_1 (B^2) \to \pi_1(S^1)
    \end{equation}
    and the composite is identity map. But observe that $B^2$ is a
    convex set and hence its fundamental group is trivial. But $S^1$
    has non-trivial fundamental group. It is then impossible to form
    such a chain of homomorphism whose product is identity map.
    Contradiction!
\end{proof}
\begin{remark}
    This theorem can be extended to higher dimensional case. But the
    proof cannot be the same because for higher dimension $\pi_1(S^n)$
    is no longer non-trivial.
\end{remark}

@@

Exmaples from the book:
\begin{ex}
    Homeomorphic spaces have the same homotopy type.
\end{ex}
\begin{ex}
    Any convex subset of a euclidean space is homotopy equivalent to a
    point.
\end{ex}
\begin{ex}
    $\E^n\setminus \{0\}$ has the homotopy type of $S^{n-1}$. This is
    shown on page 104 of \cite{book}, and is illustrated there by:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/En-0-and-S-n-1.PNG}
        \caption{$E^n\setminus\{0\}$ and $S^{n-1}$}
    \end{figure}
\end{ex}

>>>>

\paragraph{The $n$-sphere $S^n$} is on page 99. To prove it, we
require a theorem:
\begin{thm}
    Let $X$ be a space which can be written as the union of two simply
    connected open sets $U,V$ in such a way that $U\cap V$ is
    path-connected. Then $X$ is simply connected.
\end{thm}
\begin{proof}
    Proof is on page 99 of \cite{book}. It requires the Lebesgue lemma
    \ref{lemma:lebesgue-lemma}. A picture illustrating the proof is
    provided here:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]
        {pics/n-sphere-is-simply-con.PNG}
    \end{figure}
\end{proof}
Then, the two cover of $S^n$ are simply $U=S^n-\{x\}$ and
$V=S^n-\{y\}$, where $x,y$ are two distinct points on $S^n$. Since $U$
and $V$ are homeomorphic to $\E^n$, $S^n$ is simply-connected.

@@

We have an application,
\begin{thm}[Brouwer Fixed Point theorem]
    \label{thm:brouwer-fixpoint-B2}
    A contiuous map $f:B^2\to B^2$ ($B^2:2D$-closed dicks) must have a
fixed point. That is, $\exists x\in B^2$ such that $f(x)=x$.
\end{thm}
\begin{proof}
    Assuming that this theorem is false, that is $\forall x\in B^2$,
    $x\neq f(x)$, then we have a straight path from $f(x)$ to $x$. We can
    extends this path to cuts the boundary of $B^2$ at $h(x)$. This is
    for all $x\in B^2$, hence we have a map $h: B^2 \to S^1$. Also,
    $h|_{S^1}$ is obvious an identity map. But $S^1$ can be included
    inside $B^2$, so we have:
    \begin{equation}
        S^1 \to B^2 \to S^1
    \end{equation}
    Hence we have a series of homomorphism of fundamental groups:
    \begin{equation}
        \pi_1 (S^1) \to \pi_1 (B^2) \to \pi_1(S^1)
    \end{equation}
    and the composite is identity map. But observe that $B^2$ is a
    convex set and hence its fundamental group is trivial. But $S^1$
    has non-trivial fundamental group. It is then impossible to form
    such a chain of homomorphism whose product is identity map.
    Contradiction!
\end{proof}
\begin{remark}
    This theorem can be extended to higher dimensional case. But the
    proof cannot be the same because for higher dimension $\pi_1(S^n)$
    is no longer non-trivial.
\end{remark}

@@

Exmaples from the book:
\begin{ex}
    Homeomorphic spaces have the same homotopy type.
\end{ex}
\begin{ex}
    Any convex subset of a euclidean space is homotopy equivalent to a
    point.
\end{ex}
\begin{ex}
    \label{ex:Euclidean-space-and-sphere-homotopy}
    $\E^n\setminus \{0\}$ has the homotopy type of $S^{n-1}$. This is
    shown on page 104 of \cite{book}, and is illustrated there by:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/En-0-and-S-n-1.PNG}
        \caption{$E^n\setminus\{0\}$ and $S^{n-1}$}
    \end{figure}
\end{ex}

### 2016-12-27 Basic Topology Notes/sections/ch8-Simplicial-Homology.tex
[major] Talks about stellar subdivision and how homology group are a homotopic invariant.

\subsection{Section 8.4 Simplicial maps}
\label{sec:Simplicial maps}

@@

>>>>

\subsection{Section 8.4 Simplicial maps}
\label{sec:Simplicial-maps}

@@

\subsection{Section 8.5 Stellar Subdivision and Section 8.6 Invariance}
\label{sec:Stellar-Subdivision-Invariance}

Let me draw a \textbf{general blueprint}. We aims to
show that simplicial homology groups are actually a topological
property. The ultimate aim is to prove these theorems

\begin{thm}
    Any continuous function $f:|K|\to |L|$ induces a homomorphism
    $f_*:H_q(K)\to H_q(L)$ in for each $q \geq 0$.
\end{thm}

\begin{thm}
    If $f$ is the identity map of $|K|$, then each 
    $f_*:H_q(K)\to H_q(K)$ is the identity homomorphism, and if we
    have two maps
    $$ \begin{tikzcd}[]
        \lvert K\rvert\ar[r,"f"] & \lvert L\rvert\ar[r,"g"] 
        & \lvert M\rvert
    \end{tikzcd}$$
    then $(g\circ f)_* = g_*\circ f_*: H_q(K)\to H_q(M)$
    for all $q\geq 0$.
\end{thm}

\begin{thm}
    If $f,g:|K|\to |L|$ are homotopic maps, then $f_*=g_*:H_q(K)\to
    H_q(L)$ for all $q$.
\end{thm}

\begin{coro}
    If $|K|$ and $|L|$ have the same homotopy type, then $K$ and $L$
    have isomorphic homology groups.
\end{coro}

These theorems can be found on page 189 of the book\cite{book}. But
unfortunately, this course is too limited to talk about their proof in
detail. So here we provide the general picture. The starting point
will be of course, approximating a continuous function $f$ with a
simplicial map $s$. But in doing so, we have to do barycentric
subdivision on the codomain. So if we can ensure that the homology
groups are unaltered in barycentric subdivision, then of course
simplicial maps will preserve homology groups, as we have seen in
section~\ref{sec:Simplicial-maps}. And it is nature, though it
requires proof, that the function $f$ approximated by $s$ will also
have the same property of preserving simplicial maps.

This section (section 8.5) deals with proving that the barycentric
subdivision does not alter Homology groups. But it uses a simpler
step. In general, a lot more vertices will be introduced in one
barycentric subdivision. But we can look at a simper notion first,
i.e. a \nomen{stellar subdivision}. Stellar meanings relating to
stars, and a stellar subdivision quite resembles a star's shining
lines. A first example is given on top of page 186 of \cite{book}.
Here I reproduced the figure 8.6 in a more "star" like depiction:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{pics/ch8/stellar-subd-ex.pdf}
\end{figure}

The benefit of stellar subdivision is that one can do a complete
barycentric subdivision through performing a series of steller subdivision
in a specific order. Amazingly the book\cite{book} do not provide a
proof of this, except that it gives an example using figure 8.7 on
page 186. But the proof should be easy.

I have omitted a precise description of stellar subdivision here,
because it is best understood along with the process to obtain a
barycentric subdivision through stellar subdivision. They are found in
page 186 of \cite{book}. Just keep in mind that $A_0<A_1<\cdots<A_k$,
for a simplex $\{\hat{A}_0,\hat{A}_1,\cdots,\hat{A}_k\}$ in a
barycentric subdivision.

Then the book proves that
\begin{thm}
    If $K'$ is obtained from $K$ by a single stellar subdivision, then
    $K'$ and $K$ have isomorphic Homology groups
\end{thm}
and hence
\begin{coro}
    Barycentric subdivision does not change the Homology groups of a
    complex
\end{coro}

The proof can be in page 186 to 188 of \cite{book}. I also omit the
proof here, but I warn anyone who wish to study the prooves in the
next section to study the proofs here carefully, because a important
map (subdivision chain map) will be constructed here.

As an effort to prevent this tacit talk from being a transient memory,
I will mention several applications.

\begin{remark}
    First, Homology group is a homotopic property. This means we can
    calculate the Homology group of a big complex space by shrinking
    it into a simple space.
\end{remark}

\begin{key}
    The Homology groups of $S^n$ are:
    $$ H_0(S^n)=H_n(S^n)=\Z $$
    $$ H_q(S^n)=0,\text{for $q\neq 0,n$} $$
\end{key}
\begin{thm}
    $S^m$ is not homotopic to $S^n$ if $m\neq n$.
\end{thm}
\begin{proof}
    Obviously.
\end{proof}

\begin{thm}
    Two Euclidean spaces are homeomorphic if and only if they have the
    same dimension
\end{thm}
\begin{proof}
    As mentioned in
    example~\ref{ex:Euclidean-space-and-sphere-homotopy}, the 
    homotype type of $S^{m-1}$ and $\E^{m}\setminus\{0\}$ are the
    same. If there is a homotopic equivalence from $\E^m\to \E^n$, it
    would induces homotopic equivalence on $S^m$ and $S^n$. Hence
    $\E^m$ is homotopic to $\E^n$ if and only if $m=n$.
\end{proof}

\begin{thm}[Brouwer fixed-point theorem]
    A map from $B^n$ to itself must leave at least one point fixed.
\end{thm}
\begin{proof}
    The proof is exactly as it is in
    theorem~\ref{thm:brouwer-fixpoint-B2}, except that we have to use
    Homology groups instead of the fundamental group to distinguish
    different spheres.
\end{proof}
